Measuring Interaction Design \\before Building the System: a Model-Based Approach

### 2016-01-14 paper.tex
revised background and discussion sections.
According to what CHI reviewrs said and considering that EICS is more
focussed on model-based approaches.

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={SIGCHI Conference Proceedings Format},
  pdfauthor={LaTeX},
  pdfkeywords={SIGCHI, proceedings, archival format},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
}

@@

  We present a tool (\migtool) that transforms models of a user
  interface into a graph, upon which usage scenarios can be easily
  specified, and used to compute possible interaction paths. Metrics
  based on possible paths, with or without user errors, can be easily
  computed.  For example, when applying them to compare mail
  applications, we show that Gmail has 3 times more shortest routes,
  has 2x more routes that include a single user error, has routes with
  13\% fewer steps, but optimal routes have the smallest probability
  to be chosen.

@@

The literature on using state-transition networks for specifying or
analyzing the behavior of UIs is vast. We conducted a systematic-style
literature review, using Google Scholar and queries with combinations
of these phrases: ``user interface'', ``path analysis'', ``user
trace'', ``interaction trace'', ``navigation'', ``markov chain'',
``markov model'', ``state transition'', ``statechart'', ``metric'',
``measure''. For each query we analyzed title and abstract of the
first 50 hits and appraised their relevance based on whether the paper
discusses approaches for measuring user interfaces in the context of
usability and on the basis of a state transition model. This resulted
in 78 full-text papers that were later on re-analyzed against the same
criterion, leading to several of papers mentioned below.
%
Many papers deal with testing user interfaces, and problems related to
generating execution traces as a means to assess coverage of a test
suite~\cite{ammann2008introduction,briand2005:test:uml}. They were
excluded from the review.

@@

In many cases \emph{statecharts} are used, a generalization of finite
state automata.  Horrocks showed how statecharts can be used to model
and specify the dynamics of typical desktop
UIs~\cite{horrocks99}. While providing many interesting insights on
how and why one should use statecharts to do so, this nice work does
not address how such a specification could be \emph{automatically}
processed.
%
This idea was later on expanded by Thimbleby~\cite{thimbleby07};
statecharts are seen as crucial representations that allow a designer to
fully appreciate how devices behave. The overall claim is that ``If
you don't understand the logic conveyed by a statechart model of a
user interface, then you don't understand the behavior of the user
interface''.

@@

In~\cite{conversy07} the Interactive Cooperative Objects formalism is
discussed, which is based on Petri nets, yet another state transition
formalism.  It is used to enrich the ARINC 661 specification of
Cockpit Display System so that the semantics of widgets can be
expressed and the behavior of a UI be formally analyzed.

%  that are now under provides precise information
% for communication protocol between application (called User
% Applications) and user interface components (called widgets) as well
% as precise information about the widgets themselves. However, in ARINC
% 661, no information is given about the behaviour of these widgets and
% about the behaviour of an application made up of a set of such
% widgets.  The Interactive Cooperative Objects (ICOs) formalism is a
% formal description technique dedicated to the specification of
% interactive systems [4, 11]. It uses concepts borrowed from the
% object-oriented approach to describe the structural or static aspects
% of systems, and uses high-level Petri nets [8] to describe their
% dynamic or behavioural aspects.

Finite state representations have been used also as a conceptual
framework for writing the code of widgets so that events and event
handlers in the UI can more easily be conceived, developed and
verified (\eg, \cite{appert2008swingstates}).

@@

A review of usability measuring practices~\cite{hornbaek2006current}
lists, under the headings \emph{measures of efficiency/usage
  patterns}, number of keystrokes, mouse clicks, and visited objects
as possible metrics. The notion of \emph{deviation from optimal
  solution} is discussed only in the context of tools for route
planning in 3D navigation. In this paper we provide our own definition
of deviation, which applies to interaction with the UI; we provide
also our notion of execution traces across states of the UI, and the
length of these traces can be used as a measure of efficiency.

@@

A usability analysis method capable to analytically predict task
completion times from a storyboard of the UI is
CogTool~\cite{bellamy2011cogtool}. CogTool relies on the ACT-R
cognitive modeling engine, and allows a designer to setup a
low-fidelity prototype of the UI. After deciding which interaction
modality and which widgets are used to implement actions, the designer
gets an estimation of how long an experienced user would
spend on each step. CogTool takes care of adding extra ``mental''
steps, such as ``think'' steps, before certain patterns of provided
steps, according to the cognitive theory underlying ACT-R. As a
result, users of CogTool obtain the breakdown of the times required by
each of the steps. Our method is weaker in terms of cognitive validity
and in terms of precision of the output: it does not provide expected
completion times. However, with our method one can analyze a large
part of the UI, get information about possible problems in some areas,
and only then devolve more resources in building storyboards and
in making assumptions regarding widgets so that specific execution
paths previously identified can be analyzed with CogTool. In a sense,
the output of our method could be used to make informed decisions as
to what to analyze with CogTool.

@@

\migtool\ takes as input UML statecharts, which are a generalization
of finite state automata (FSA), with an expressive language that
includes hierarchic levels of abstraction, concurrent regions, states
and pseudostates, guards, and an extended state notion based on an
arbitrary underlying computational model.  Harel, the inventor of
statecharts, gives an interesting retrospective view of how they were
invented, back in early eighties, and why they were appealing also to
non experts~\cite{harel09}.  For the sake of brevity, we refrain from
describing them here in detail; the interested reader is referred to
the UML standard~\cite{uml25} or some of the textbooks that deal with
them, like~\cite{samek09}.

@@

\subsection{Defining usage scenarios}

@@

Models of the four applications have been manually defined. In order
to support a fair comparison, all four models cover the same set of
use cases: listing the content of the inbox, reading a message or
conversation, replying to a message, composing and sending a new
message.

@@

The model that we show here is part of what we used in the examples
reported below. The actual model that we used consists of 
24 states, 12 pseudo states, 12 regions, and 61 transitions.

@@

Inspection of the interaction graph is not particularly useful because
even for small graphs like the one obtained from our Gmail model no
particular structure is evident that was not known from the
model. Figure~\ref{fig:gom} shows a plot using a circular layout of
the 47 states: because of the large number of cycles that exist among
states, edges form an intricate web of possible action sequences. We
believe this in practice greatly reduces usefulness of typical network
analysis metrics, such as \emph{betweenness, eccentricity, page-rank,
  eigenvalue} centrality measures.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figures/gom.png}
  \caption{Interaction graph obtained from the Gmail model.}
  \label{fig:gom}
\end{figure}

@@

Execution traces can be used also to assess what is the effect of
adding a widget or feature to an existing UI.  For example, we studied
some cruise control features of cars. One of the examples is system
S\footnote{There is no need to disclose the actual brand.}, where the
driver can engage the system, and once it is engaged, speed can be
increased or decreased with small or large steps. Of course the system
can also be disengaged (by pressing the brake pedal, for example).
System A is more elaborate, as it includes also a memory function:
when it is disengaged it remembers the current speed, which can be
recalled later on.  There are two ways to re-engage it: one by setting
a new speed, and one by recalling the previous one. In addition, if
the car drives for more than 5 minutes at a higher speed than the set
one, system A automatically disengages.

@@

When using a scenario for S where we assume that re-setting the speed
is done manually by the driver with 4 actions on ``up'' and ``down''
to approximately get the same speed), the comparison produces the
values shown in Table~\ref{tab:data}. 
System A has 2 alternative optimal paths, their length is 4.5 (36\%
shorter than in S); both systems have a comparable number of detour 1
traces, but again their length in A is  28\% shorter. The probability
of hitting the optimal trace is however twice as large in S, which
features also a more compact design. 
In both cases detours are caused by the possibility of disengaging the
system at the wrong moment.

Therefore we can conclude that:
\begin{inparaenum}[(1)]
\item system A makes users more efficient (an increase of 56\%) for
  the considered scenario;
\item with A there are two possible ways to achieve the scenario, thus
  more flexibility is given;
\item with A the probability of doing the right thing is almost half
  of system S: it might be more difficult to do the right thing
  because more possibilities are offered;
\item system A features a more compact design, with fewer action types
  to be performed at each moment.
\end{inparaenum}

@@

Expressivity of the modeling language means that different designers
are likely to produce different models for the same UI. As a
consequence, it is possible that metrics computed by \migtool\ do depend on different
modeling choices. It is worth mentioning, though, that because of the
flattening process, several differences are reduced (for example,
those dealing with using a different hierarchy of states, or with
distributing differently concurrent regions across states), and 
sensitivity is correspondingly reduced. At the moment, however, we do
not have hard evidence to back this claim.

Differently from other model-based approaches, such as those based on
IFML, \migtool\ uses \emph{only} a model of the behavior of the
UI. Designers do not need to cope with data modeling, nor with
decisions dealing with presentation. In a sense, \migtool\ uses only the
\emph{controller} part of the Model-View-Controller paradigm that is
adopted when developing UIs. Therefore, the designer using \migtool\ to
perform analysis is free from other concerns that in the end affect
usability, and the conclusions that are derived with \migtool\ can be
combined with other results \emph{after} the analysis is
performed.

In terms of validity of conclusions obtained through \migtool, we can
say that because they are devoid of user behavior assumptions (such as
preferences, skills, interpretations, ergonomic constraints) they are
very general and conservative. On the other hand, predictions based on
\migtool\ metrics are also generic because they do not consider data
and presentation aspects, nor user-related factors. For example, it is
unfeasible to use \migtool\ to predict the time needed by a user to
complete a scenario. However, as mentioned before, \migtool\ can be
used to analyze the whole interaction design and gather data to inform
more specific analyses that could be performed, for example, with
CogTool.

@@

Likewise it is possible to extend \migtool\ to compute also the
lostness metric~\cite{smith1996lostness}, based on states with
different detour orders. This would give yet another objective metric
to compare different designs. We opted for non doing it in light of
the fact that lostness was designed for exploratory activities, not
for goal oriented scenarios like the ones we considered in previous
examples.

\migtool\ is part of a suite of model-based tools for designing,
analyzing and testing user interfaces developed by the company
ANONYMIZED. \migtool\ is implemented partly in Java (model processing)
and partly in R/iGraph. Development of a web-based UI of \migtool\ is
underway; it will be freely available for research purposes.

@@

% REFERENCES FORMAT
% References must be the same font size as other body text.
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{gb,database}

>>>>

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={Measuring Interaction Design},
  pdfauthor={LaTeX},
  pdfkeywords={model-based User Interfaces},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
}

@@

  We present a tool (MIGTool) that transforms models of the behavior
  of a user interface into a graph, upon which usage scenarios can be
  easily specified, and used by MIGTool to compute possible
  interaction paths. Metrics based on possible paths, with or without
  user navigation errors, can then be computed.  For example, when
  analyzing four mail applications, we show that Gmail compared to
  others has 3 times more shortest routes, has 2x more routes that
  include a single user error, has routes with 13\% fewer steps, but
  has also optimal routes with the smallest probability to be chosen.

  Without MIGTool, this kind of analysis could only be done after
  building some prototype of the system, and then only for specific
  scenarios by manually tracing user actions and relative changes to
  the screens. With MIGTool the exploration of suitability of a design
  with respect to different scenarios, or comparison of different
  design alternatives, can be done after providing only a partial
  specification of the user interface behavior.

  This is made possible by the ability to associate scenarios steps to
  required user actions as defined in the model, by an efficient
  strategy to identify complete execution traces that users can
  follow, and by computing a range of diverse metrics on these
  results. The agile modeling approach that was followed implies that
  relatively little work is needed upfront.

  % CONTRIBUTION (asked when submitting the abstract)

  % There are three contributions.

  % First, we provide means to concisely ground usage scenarios on the
  % basis of actions specified in models. This mechanism is general
  % enough that can be used also to link tasks or use cases to actual
  % behavior of a user interface.

  % Second, MIGTool transforms models and scenarios into a graph of
  % possible execution paths. Thanks to the special-purpose language
  % that is provided to specify scenarios, MIGTool is quite efficient,
  % with a time complexiy that is linear with the number of steps in a
  % scenario and the number of edges in the graph. We describe the
  % language and the algorithm.

  % Third, we provide definition of a handful of metrics that are
  % objective, precise and that bear upon usability of the resulting
  % user interface. They are very general, can be used for most kinds of
  % user interfaces that can be modeled with discrete events, and are
  % platform independent.

%% EICS submission: #122
%% https://precisionconference.com/~eics16a/updatePaperResponse?userRef=CILXDODXPNF0VESOAFMMC8I16ZO5FR5J&paperNumber=112&emailSent=&error=

@@

The literature on using state-transition networks for specifying or
analyzing the behavior of UIs is vast. We conducted a systematic-style
literature review, using Google Scholar and queries with combinations
of these phrases: ``user interface'', ``path analysis'', ``user
trace'', ``interaction trace'', ``navigation'', ``markov chain'',
``markov model'', ``state transition'', ``statechart'', ``metric'',
``measure''. For each query we analyzed title and abstract of the
first 50 hits and appraised their relevance based on whether the paper
discusses approaches for measuring user interfaces in the context of
usability and on the basis of a state transition model. This resulted
in 78 full-text papers that were later on re-analyzed against the same
criterion, leading to several of papers mentioned below.
%
% Many papers deal with testing user interfaces, and problems related to
% generating execution traces as a means to assess coverage of a test
% suite~\cite{ammann2008introduction,briand2005:test:uml}. They were
% excluded from the review.

@@

In many cases \emph{statecharts} are used, a generalization of finite
state automata.  Horrocks showed how statecharts can be used to model
and specify the dynamics of typical desktop
UIs~\cite{horrocks99}. While providing many interesting insights on
how and why one should use statecharts to do so, this nice work does
not address how such a specification could be \emph{automatically}
processed.
%
This idea was later on expanded by Thimbleby~\cite{thimbleby07};
statecharts are seen as crucial representations that allow a designer to
fully appreciate how devices behave. The overall stance is that ``If
you don't understand the logic conveyed by a statechart model of a
user interface, then you don't understand the behavior of the user
interface''.

@@

% In~\cite{conversy07} the Interactive Cooperative Objects formalism is
% discussed, which is based on Petri nets, yet another state transition
% formalism.  It is used to enrich the ARINC 661 specification of
% Cockpit Display System so that the semantics of widgets can be
% expressed and the behavior of a UI be formally analyzed.
% 
%  that are now under provides precise information
% for communication protocol between application (called User
% Applications) and user interface components (called widgets) as well
% as precise information about the widgets themselves. However, in ARINC
% 661, no information is given about the behaviour of these widgets and
% about the behaviour of an application made up of a set of such
% widgets.  The Interactive Cooperative Objects (ICOs) formalism is a
% formal description technique dedicated to the specification of
% interactive systems [4, 11]. It uses concepts borrowed from the
% object-oriented approach to describe the structural or static aspects
% of systems, and uses high-level Petri nets [8] to describe their
% dynamic or behavioural aspects.

Finite state representations have been used also as a conceptual
framework for writing the code of widgets so that events and event
handlers in the UI can more easily be conceived, developed and
verified (\eg, \cite{appert2008swingstates}). More sophisticated
model-based approaches employ statecharts to orchestrate the behavior of
different components of a user interface \cite{winckler03}. These
models, however, are not used to support usability analysis of any
kind.

@@

@@

The notion of \emph{deviation from optimal
  solution} is discussed only in the context of tools for route
planning in 3D navigation. In this paper we provide our own definition
of deviation, which applies to interaction with the UI; we provide
also our notion of \emph{potential execution traces} across states of
the UI, and the length of these traces can be used as a measure of
efficiency.

In \cite{fayollas2014}, hierarchical task models (specified with
HAMSTERS) are used to specify how pilots are expected to interact with
airplane cockpit interfaces in tasks such as changing value to some
flight parameter. The paper in particular discusses how these task
models can be used to specify alternative ways to achieve te same
goal, and analyse the consequences on the reliability of the resulting
system. Compared to MIGTool, that approach cannot take advantage of a
model of the system and has instead to rely only on models of the
tasks. The conseguence is that analysing alternative ways to achieve
the same goal requires developing different models; in addition,
exploring suitability of a given system with respect to certain tasks
is complex; and finally, hierarchical task models cannot take
advantage of contexts and rich control flows that are available with
state-transition representations, limiting therefore the richness of
the scenarios being analysed.

A review of usability measuring practices~\cite{hornbaek2006current}
lists, under the headings \emph{measures of efficiency/usage
  patterns}, number of keystrokes, mouse clicks, and visited objects
as possible metrics. Notice that in general the time needed to carry
out individual interaction steps depends on many aspects, such as the
time to decide what goal to achieve, to understand how to do it, to
locate the actions to do it, to perform the action, to attend to the
feedback. Layout of the user interface, Fitt's law considerations,
cognitive state of the user, physical and environmental aspects affect
such times. For these reasons we opted, in \migtool, to only focus on
required steps to complete a task, not on completion time.

A usability analysis method capable to analytically predict task
completion times from a storyboard of the UI is
CogTool~\cite{bellamy2011cogtool}. CogTool relies on the ACT-R
cognitive modeling engine, and allows a designer to setup a
low-fidelity prototype of the UI. After deciding which interaction
modality and which widgets are used to implement actions, the designer
gets an estimation of how long an experienced user would spend on each
step. CogTool takes care of adding extra ``mental'' steps, such as
``think'' steps, before certain patterns of provided steps, according
to the cognitive theory underlying ACT-R. As a result, users of
CogTool obtain the breakdown of the times required by each of the
steps. Our method is less precise: it does not provide expected
completion times that are cognitively justified. However, with our
method, you can analyze a large part of the UI, get information about
possible problems in some areas, and only then devolve more resources
in building storyboards and in making assumptions regarding widgets so
that specific execution paths previously identified can be analyzed
with CogTool. In a sense, the output obtained with our method could be
used to make informed decisions as to what to analyze next with
CogTool. Approaches like \cite{swearngin13} start from the assumption
that the user interface exists already, and that it can \emph{ripped}
to reverse engineer a behavior model. These assumptions are not needed
to use \migtool.

Another approach for analysing user actions that has a strong
cognitive background is SNIF-ACT \cite{pirolli2007}. However it has
been conceived and applied to information websites, where the notion
of \emph{information scent} can bear upon interaction. We opted for
not applying such concepts to the interaction with a web application,
where most of the actions are not aimed at information finding.

@@

\migtool\ takes as input UML statecharts, which are a generalization
of finite state automata (FSA), with an expressive language that
includes hierarchic levels of abstraction, concurrent regions, states
and pseudostates, guards, and an extended state notion based on an
arbitrary underlying computational model.  Harel, the inventor of
statecharts, gives an interesting retrospective view of how they were
invented, back in early eighties, and why they were appealing also to
non experts~\cite{harel09}.%   For the sake of brevity, we refrain from
% describing them here in detail; the interested reader is referred to
% the UML standard~\cite{uml25} or some of the textbooks that deal with
% them, like~\cite{samek09}.

@@

\subsection{Grounding usage scenarios}

@@

Models of the four applications have been manually defined. In order
to support a fair comparison, all four models cover the same set of
use cases at the same level of detail: listing the content of the
inbox, reading a message or conversation, replying to a message,
composing and sending a new message.

@@

The model that we show here is part of what we used in the examples
reported below. The actual model that we used consists of 
24 states, 12 pseudo states, 12 regions, and 61 transitions. The model
is similar to that being discussed in examples mentioned by \cite{ifml13}.

@@

Inspection of the interaction graph is not particularly useful because
even for small graphs like the one obtained from our Gmail model no
particular structure is evident that was not known from the
model.
%  Figure~\ref{fig:gom} shows a plot using a circular layout of
% the 47 states: 
Because of the large number of cycles that exist among states, edges
in the interaction graph form an intricate web of possible action
sequences. We believe this in practice greatly reduces usefulness of
typical network analysis metrics, such as \emph{betweenness,
  eccentricity, page-rank, eigenvalue} centrality measures.

% \begin{figure}
%   \centering
%   \includegraphics[width=\linewidth]{figures/gom.png}
%   \caption{Interaction graph obtained from the Gmail model.}
%   \label{fig:gom}
% \end{figure}

@@

Execution traces can be used also to assess what is the effect of
adding a widget or feature to an existing UI.  For example, we studied
some cruise control features of cars. One of the examples is system
S\footnote{There is no need to disclose the actual brand.}, where the
driver can engage the system, and once it is engaged, set speed can be
increased or decreased with small or large steps. Of course the system
can also be disengaged (by pressing the brake pedal, for example).
System A is more elaborate, as it includes also a memory function:
when it is disengaged it remembers the currently set speed, which can be
recalled later on.  There are two ways to re-engage it: one by setting
a new speed, and one by recalling the previous one. In addition, if
the car drives for more than 5 minutes at a higher speed than the set
one, system A automatically disengages.

@@

When using a scenario for S where we assume that re-setting the speed
is done manually by the driver with 4 actions on ``up'' and ``down''
to approximately get the same speed), the comparison produces the
values shown in Table~\ref{tab:data}.  System A has 2 alternative
optimal paths, their length is 4.5 (36\% shorter than in S); both
systems have a comparable number of detour 1 traces ($N_1$), but again
their length in A ($L_1$) is 28\% shorter. The probability of hitting
the optimal trace is however twice as large in S, which features also
a more compact design.  In both cases detours are caused by the
possibility of disengaging the system at the wrong moment.

Therefore we can conclude that:
\begin{inparaenum}[(1)]
\item system A makes users more efficient (an increase of 56\%) for
  the considered scenario;
\item with A there are two possible ways to achieve the scenario, thus
  more flexibility is given;
\item with A the probability of doing the right thing is almost half
  of system S: it might be more difficult to do the right thing
  because more possibilities are offered;
\item system S features a more compact design, with fewer action types
  to be performed at each moment.
\end{inparaenum}

@@

Expressivity of the modeling language means that different designers
are likely to produce different models for the same UI. As a
consequence, it is possible that metrics computed by \migtool\ do
depend on different modeling choices. It is worth mentioning, though,
that because of the flattening process, several differences are
reduced (for example, those dealing with using a different hierarchy
of states, or with distributing differently concurrent regions across
states), and sensitivity is correspondingly reduced. % At the moment,
% however, we do not have hard evidence to back this claim.
Future studies for addressing this issue could follow what was done in
\cite{john10}.

Differently from other model-based approaches, such as those based on
IFML, \migtool\ uses \emph{only} a model of the behavior of the UI. In
IFML the modeler specifies behavior in terms of concepts that are very
close to statecharts (IFML adopts a dual representation of
AND/OR-states, has guards, has events, has an underlying event
broadcasting mechanism) upon which data binding is applied, so that
transitions can be easily bound to data instances. On the other hand,
designers using \migtool\ do not need to cope with data modeling, nor
with decisions dealing with presentation. In a sense, \migtool\ uses
only the \emph{controller} part of the Model-View-Controller paradigm
that is adopted when developing UIs. Although in principle the same
results could be obtained from IFML models, to do so one needs to
isolate the control part of the IFML model from the data part; this is
difficult to do because, by design, IFML integrates very well the two
aspects.

A designer using \migtool\ is free from other concerns that in the end
affect usability, and the conclusions that are derived with \migtool\
can be combined with other results \emph{after} the analysis is
performed. As mentioned above, ours is an agile modeling approach
\cite{ambler02} that:
\begin{inparaenum}[(a)]
\item enables analysis of a UI well before it is developed;
\item  requires \emph{only} a model of its behavior;
\item  supports exploration of different tasks/scenarios;
\item enables grounding of tasks/scenarios through a
  platform-independent specification. As such, grounding can be viewed
  as the specification of a \emph{reification} relation between a
  \emph{task-and-domain-concepts} level to an
  \emph{abstract-user-interface} level (we are referring to the
  CAMELEON reference model \cite{calvary:iwc03:camaleon}).
\end{inparaenum}
It is possible to extend \migtool\ so that also task models
(specified, for example, through CTT, UsiXML, HAMSTERS
\cite{mori02,limbourg04:usixml,martinie15}) can be processed and used
as specification of the scenarios to be analyzed. At that point,
\migtool\ could be used as a tool to explore suitability of tasks with
respect to a behavioral model of a user interface. This problem has
been addressed in \cite{martinie15} through a mechanism that relates
task models to source code annotations of the implementation of the
user interface. In this way it is possible to simulate execution of
the tasks in parallel with execution of the system and ensure
compatibility of the system with respect to tasks. With an extended
version of \migtool\ it could be done without having to implement the
system.

In terms of validity of conclusions obtained through \migtool, we can
say that because they are devoid of user behavior assumptions (such as
preferences, skills, interpretations, ergonomic constraints) they are
very general and conservative. On the other hand, predictions based on
\migtool\ metrics are also generic because they do not consider data
and presentation aspects, nor user-related factors. For example, it is
unfeasible to use \migtool\ to predict the time needed by a user to
complete a scenario. However, as mentioned before, \migtool\ can be
used to analyze the whole interaction design and gather data to inform
more specific analyses that could be performed, for example, with
CogTool or CogTool-Helper \cite{swearngin13}.

@@

Likewise it is possible to extend \migtool\ to compute also the
lostness metric~\cite{smith1996lostness}, based on states with
different detour orders. This would give yet another objective metric
to compare different designs. We opted for non doing it in light of
the fact that lostness was designed for exploratory activities in
information websites, not for goal oriented scenarios like the ones we
considered in previous examples.

% \migtool\ is part of a suite of model-based tools for designing,
% analyzing and testing user interfaces developed by the company
% ANONYMIZED. 
\migtool\ is implemented partly in Java (model processing)
and partly in R/iGraph. Development of a web-based UI of \migtool\ is
underway; it will be freely available for research purposes. Currenlty
\migtool\ reads UML models represented as XMI files; it could be
easily extended to handle also SCXML representations of statecharts \cite{scxml}.

@@

\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{gb,database}
